---
title: "Coding Simulations"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{coding-simulations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup}
library(ordinalsimr)
```


# Coding Your Own Simulations

This guide will provide a rough overview of how to code your own simulations using the components of this package should you find the Shiny application too limiting for your own purposes. Key information on functions:

* `run_simulations()` will take simulation input parameters and return a list. The first element contains a matrix of p-values for each run of the simulation, while the second element is multiple lists of information describing group allocations in detail.
* `format_simulation_data()` is recommended for reformatting the two list elements mentioned above. It will return a list of two tibbles.
* `calculate_power_t2error()` and `calculate_t1_error()` can receive the p-value data frames for performing T1 Error, Power, and T2 Error calculations with confidence intervals. See function documentation for additional arguments.

## Power and Type II Error

```{r, warning=FALSE}
sim_results <- run_simulations(
  sample_size = 80,
  sample_prob = c(0.5, 0.5),
  prob0 = c(0.1, 0.2, 0.3, 0.4),
  prob1 = c(0.6, 0.2, 0.1, 0.1),
  niter = 20
)

formatted_results <- format_simulation_data(sim_results)
names(formatted_results)
head(formatted_results$p_values)

formatted_results %>%
  select(Wilcoxon, Fisher, `Chi Squared\n(No Correction)`, 
         `Chi Squared\n(Correction)`, `Prop. Odds`, 
         `Coin Indep. Test`,
         sample_size) %>%
  calculate_power_t2error(alpha = 0.05, power_confidence_int = 95)
```


## Type I Error

To find the Type I error of a distribution, the code from before is largely unchanged except for the fact that the probability vectors set in `run_simulations` must now be equivalent and the `calculate_t1_error()` function is now applied.


```{r, warning=FALSE}
sim_results <- run_simulations(
  sample_size = 30:35,
  sample_prob = c(0.5, 0.5),
  prob0 = c(.4, .3, .3),
  prob1 = c(.8, .1, .1), # note the matching probabilities between groups
  niter = 50
)

formatted_results <- format_simulation_data(sim_results)
names(formatted_results)
head(formatted_results$p_values)


formatted_results %>%
  select(Wilcoxon, Fisher, `Chi Squared\n(No Correction)`, 
         `Chi Squared\n(Correction)`, `Prop. Odds`, 
         `Coin Indep. Test`,
         sample_size) %>%
  calculate_t1_error(alpha = 0.05, t1_error_confidence_int = 95)
```



## Mapping Over Many Sample Sizes

The current version of the application can only accept one sample size at a time. A future iteration of the application will ideally implement this feature, but, in the meantime, there are many situations where it would be advantageous to model test results over many different sample sizes. Depending on how big the number of iterations per sample sizes (`niter`) and the actual number of sample sizes being checked, it may only be practical to do this in a parallelized manner with e.g. {furrr} or {parallel}. In any case, an example of such code is included below:


```{r, warning=FALSE}
sample_sizes <- c(30, 50, 100)

lapply(sample_sizes, function(x) {
  run_simulations(
    sample_size = x,
    sample_prob = c(0.5, 0.5),
    prob0 = c(0.1, 0.2, 0.3, 0.4),
    prob1 = c(0.6, 0.2, 0.1, 0.1),
    niter = 100
  ) %>%
    format_simulation_data() %>%
    select(Wilcoxon, Fisher, `Chi Squared\n(No Correction)`, 
           `Chi Squared\n(Correction)`, `Prop. Odds`, 
           `Coin Indep. Test`, sample_size) %>%
    calculate_power_t2error()
})
```

Equivalent code with purrr

``` r
sample_sizes %>%
  purrr::map(
    ~run_simulations(
      sample_size = .x,
      sample_prob = c(0.5, 0.5),
      prob0 = c(0.1, 0.2, 0.3, 0.4),
      prob1 = c(0.6, 0.2, 0.1, 0.1),
      niter = 100
      ) %>%
      format_simulation_data() %>%
      select(Wilcoxon, Fisher, `Chi Squared\n(No Correction)`, 
           `Chi Squared\n(Correction)`, `Prop. Odds`, 
           `Coin Indep. Test`, sample_size) %>%
      calculate_power_t2error()
)
```

## Adjust Multiple Parameters

It is perhaps more likely that analysts will want to iterate simulations over a variety of different parameters at once. The code below provides a structure for creating a combination grid based on the 5 input parameters that can be altered; this example can easily be altered to include desired parameters by replacing/removing/expanding the listed parameters. 


### Set Parameters

Note that the `prob0_list` and `prob1_list` must always be of the same length **and** the corresponding sub-elements of the list must also be of the same length. Put in terms of the application, there must be a Group 2 if there is a Group 1, and the vector representing the number of possible outcomes must be the same length for these 2 groups.

If performing simulations on one distribution to evaluate Type I error, it is only necessary to form one probability list. This object can be recycled for the probabilities of both groups.

```{r, warning=FALSE}
# Choose sample sizes
sample_size <- c(50, 100)
# Set sample distributions as a proportion c(group1, group2)
sample_prob <- list(c(0.5, 0.5), c(0.75, 0.25))
# Trial 1 has matching probabilities between the 2 groups. Trial 2 has non-matching probabilities
prob0_list <- list(trial1_group1 = c(0.1, 0.2, 0.3, 0.4), trial2_group1 = c(0.1, 0.2, 0.3, 0.4))
prob1_list <- list(trial1_group2 = c(0.1, 0.2, 0.3, 0.4), trial2_group2 = c(0.2, 0.3, 0.3, 0.2))
# Number of iterations
niter <- c(20, 100)
```

### Create Simulation Grid

Use `tidyr::expand_grid()` rather than `base:expand.grid()` because the former creates a tibble by default, and this supports the nested tibble structure which I'm relying on. (This can, of course, be approached in other ways if desired.)

```{r, warning=FALSE}
# Use tidyr::expand_grid as it creates a tibble, supporting the nested tibble structure
info_table <- tidyr::expand_grid(
  sample_size,
  sample_prob,
  prob0_list,
  prob1_list,
  niter
)

info_table
```



### Run Simulation

The example below is complete in running the simulations and calculating the Power and Type II error. However, the same code can be applied to either calculate the Type I error or use the p-values for other purposes.

```{r, warning=FALSE}
# Calculate either Power/T2 error or T1 error depending on your specific needs
many_sims <- mapply(
    ordinalsimr::run_simulations,
    sample_size = info_table$sample_size,
    sample_prob = info_table$sample_prob,
    prob0 = info_table$prob0_list,
    prob1 = info_table$prob1_list,
    niter = info_table$niter
)


length(many_sims)
many_sims[1]
```


Equivalent Code with {purrr}

``` r
info_table %>%
  purrr::pmap(
    ~ run_simulations(
      sample_size = ..1,
      sample_prob = ..2,
      prob0 = ..3,
      prob1 = ..4,
      niter = ..5
    ) %>%
      format_simulation_data() %>%
      magrittr::extract2("p_values") %>%
      calculate_power_t2error(),
    .progress = TRUE
  )
```



Note that even with relatively small sample sizes and a number of iterations 1-3 magnitudes less than normally specified for simulation studies, processing 32 different simulations took ~20-30 seconds. 



