---
title: "Using {ordinalsimr}"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using {ordinalsimr}}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup}
library(ordinalsimr)
```


# Shiny App

The {ordinalsimr} package wraps a Shiny application and supporting functions for running simulation studies on several pre-selected statistical tests applicable to ordinal data. The Shiny app is particularly suited for calculating Power and Type II error for a proposed 2-group comparison of an ordinal endpoint. Available parameters to manipulate before running the tests include:

* Number of simulations to run (integer)
* The sample size (integer)
* The distribution ratio between group 1 and group 2 (ratio of form X:X)
* Group 1 and Group 2 distribution probabilities for levels of the ordinal outcome

In addition to these parameters for running the simulation, the following can be adjusted in the Distributions page

* Filtering out the top x% of observations from view in the plot. *This does not remove or alter data from any other processes. It is only included as a convenience function for observing smaller windows of the data.*
* Setting the alpha significance level/p-value threshold
* Setting the confidence intervals for the Power/T2 Error calculations
* Setting the confidence intervals (separately) for the Type I Error in Group 1 and Group 2

Bug reports and feature requests can be submitted as issues at [https://github.com/NeuroShepherd/ordinalsimr/issues](https://github.com/NeuroShepherd/ordinalsimr/issues)

## App Options

If you find yourself consistently changing the default parameters of the application to new values manually, it is possible to instead set default values for these parameters. All available preset parameters are prefixed with `ordinalsimr.`, but this functionality has so far only been implemented for a limited number of defaults. Options can be set using the `options()` function. 

Example of changing the default number of rows to 6:

```
options(ordinalsimr.default_entry_rows = 6)
```

### Available Options

* `ordinalsimr.default_entry_rows`: single integer indicating number of rows
* `ordinalsimr.default_distributions`: data frame to use for simulations. 
  * Safety checks have not been included on this option. Input must be a 2-column data frame. Column names will be discarded. Recommend that you ensure the sum of values in each column are **near** 1 (e.g. using `dplyr::near()`).


### Planned Options

* `ordinalsimr.default_sample_size`: sample size
* `ordinalsimr.default_simulations`: number of simulation iterations
* ...

## Using Downloaded Data

Data downloaded after running a simulation is stored as a `.rds` file, and can be loaded into your R session using `readRDS()`. The data is structured as a named list with 3 elements at the top level, and several sub-elements. A summary of the available information is available in the code below.

```{r}
output <- readRDS("data-2024-02-16-049dcfcd9afdad1402588ca3a3f73f95-1.rds")
output$comparison_data$distribution_statistics
str(output, max.level = 2)
```

* Top level items
  * `comparison_data` are results from comparing Group 1 against Group 2 in the statistical tests (for TII error and power)
  * `group1_data` are results from comparing Group 1 data against itself (for TI error of this group)
  * `group2_data` are results from comparing Group 2 data against itself (for TI error of this group)
* Second level items
  * `p_values` are tables of p-values for all of the tests (columns) in each simulation runs (rows)
  * `run_info` are tables of detailed metainformation about the parameters used for each run
  * `distribution_statistics` are tables of computed TII error, power, and associated confidence intervals
  * `distribution_plot` is a {ggplot2} plot of the distributions of p-values
  * `group1_t1error` and `group2_t1error` are tables of TI error and associated confidence intervals



