---
title: "Using {ordinalsimr}"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using {ordinalsimr}}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ordinalsimr)
```


# Shiny App

* Brief description of how to use the app
* Alternative terminology
* Notes on future advances
* Link to bug reports

## App Options

* `ordinalsimr.default_entry_rows` for default number of rows in the data entry table on start-up
* ...

Setting options:

```
options(name = "value")
options(ordinalsimr.default_entry_rows = 6)
```

# Using Downloaded Data

* Description of the output data
* Accessing the elements of the lists


# Coding Your Own Simulations

* Describing some of the functions in more detail
* Showing example output
* Work through an entire example

## Mapping Over Many Sample Sizes

* Note this is a planned future feature for the app
* Demonstrate how to do this by using `purrr::map()`
  * recommend limiting number of runs unless using `furrr::map()` or other parallelization approach



