---
title: "Using {ordinalsimr}"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using {ordinalsimr}}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup}
library(ordinalsimr)
```


# Shiny App

The {ordinalsimr} package wraps a Shiny application and supporting functions for running simulation studies on several pre-selected statistical tests applicable to ordinal data. The Shiny app is particularly suited for calculating Power and Type II error for a proposed 2-group comparison of an ordinal endpoint. Available parameters to manipulate before running the tests include:

* Number of simulations to run (integer)
* The sample size (integer)
* The distribution ratio between group 1 and group 2 (ratio of form X:X)
* Group 1 and Group 2 distribution probabilities for levels of the ordinal outcome

In addition to these parameters for running the simulation, the following can be adjusted in the Distributions page

* Filtering out the top x% of observations from view in the plot. *This does not remove or alter data from any other processes. It is only included as a convenience function for observing smaller windows of the data.*
* Setting the alpha significance level/p-value threshold
* Setting the confidence intervals for the Power/T2 Error calculations
* Setting the confidence intervals (separately) for the Type I Error in Group 1 and Group 2

Bug reports and feature requests can be submitted as issues at [https://github.com/NeuroShepherd/ordinalsimr/issues](https://github.com/NeuroShepherd/ordinalsimr/issues)

## App Options

If you find yourself consistently changing the default parameters of the application to new values manually, it is possible to instead set default values for these parameters. All available preset parameters are prefixed with `ordinalsimr.`, but this functionality has so far only been implemented for a limited number of defaults. Options can be set using the `options()` function. 

Example of changing the default number of rows to 6:

```
options(ordinalsimr.default_entry_rows = 6)
```

### Available Options

* `ordinalsimr.default_entry_rows`: single integer indicating number of rows
* `ordinalsimr.default_distributions`: data frame to use for simulations. 
  * Safety checks have not been included on this option. Input must be a 2-column data frame. Column names will be discarded. Recommend that you ensure the sum of values in each column are **near** 1 (e.g. using `dplyr::near()`).


### Planned Options

* `ordinalsimr.default_sample_size`: sample size
* `ordinalsimr.default_simulations`: number of simulation iterations
* ...

## Using Downloaded Data

Data downloaded after running a simulation is stored as a `.rds` file, and can be loaded into your R session using `readRDS()`. The data is structured as a named list with 3 elements at the top level, and several sub-elements. A summary of the available information is available in the code below.

```{r}
output <- readRDS("data-2024-02-16-049dcfcd9afdad1402588ca3a3f73f95-1.rds")
output$comparison_data$distribution_statistics
str(output, max.level = 2)
```

* Top level items
  * `comparison_data` are results from comparing Group 1 against Group 2 in the statistical tests (for TII error and power)
  * `group1_data` are results from comparing Group 1 data against itself (for TI error of this group)
  * `group2_data` are results from comparing Group 2 data against itself (for TI error of this group)
* Second level items
  * `p_values` are tables of p-values for all of the tests (columns) in each simulation runs (rows)
  * `run_info` are tables of detailed metainformation about the parameters used for each run
  * `distribution_statistics` are tables of computed TII error, power, and associated confidence intervals
  * `distribution_plot` is a {ggplot2} plot of the distributions of p-values
  * `group1_t1error` and `group2_t1error` are tables of TI error and associated confidence intervals



# Coding Your Own Simulations

This guide will provide a rough overview of how to code your own simulations using the components of this package should you find the Shiny application too limiting for your own purposes. Key information on functions:

* `run_simulations()` will take simulation input parameters and return a list. The first element contains a matrix of p-values for each run of the simulation, while the second element is multiple lists of information describing group allocations in detail.
* `format_simulation_data()` is recommended for reformatting the two list elements mentioned above. It will return a list of two tibbles.
* `calculate_power_t2error()` and `calculate_t1_error()` can receive the p-value data frames for performing T1 Error, Power, and T2 Error calculations with confidence intervals. See function documentation for additional arguments.

## Example: Power and Type II Error

```{r, warning=FALSE}
sim_results <- run_simulations(
  sample_size = 80,
  sample_prob = c(0.5, 0.5),
  prob0 = c(0.1, 0.2, 0.3, 0.4),
  prob1 = c(0.6, 0.2, 0.1, 0.1),
  niter = 20
)

formatted_results <- format_simulation_data(sim_results)
names(formatted_results)

head(formatted_results$p_values)

calculate_power_t2error(
  formatted_results$p_values,
  alpha = 0.05,
  power_confidence_int = 95
)
```


## Example: Type I Error

To find the Type I error of a distribution, the code from before is largely unchanged except for the fact that the probability vectors set in `run_simulations` must now be equivalent and the `calculate_t1_error()` function is now applied.


```{r, warning=FALSE}
sim_results <- run_simulations(
  sample_size = 80,
  sample_prob = c(0.5, 0.5),
  prob0 = c(0.1, 0.2, 0.3, 0.4),
  prob1 = c(0.1, 0.2, 0.3, 0.4),
  niter = 20
)

formatted_results <- format_simulation_data(
  sim_results
)
names(formatted_results)

head(formatted_results$p_values)

calculate_t1_error(
  formatted_results$p_values,
  alpha = 0.05,
  t1_error_confidence_int = 95
)
```



## Example: Mapping Over Many Sample Sizes

The current version of the application can only accept one sample size at a time. A future iteration of the application will ideally implement this feature, but, in the meantime, there are many situations where it would be advantageous to model test results over many different sample sizes. Depending on how big the number of iterations per sample sizes (`niter`) and the actual number of sample sizes being checked, it may only be practical to do this in a parallelized manner with e.g. {furrr} or {parallel}. In any case, an example of such code is included below:


```{r, warning=FALSE}
sample_sizes <- c(30, 50, 100)

purrr::map(sample_sizes,
  ~ run_simulations(
    sample_size = .x,
    sample_prob = c(0.5, 0.5),
    prob0 = c(0.1, 0.2, 0.3, 0.4),
    prob1 = c(0.6, 0.2, 0.1, 0.1),
    niter = 100
  ) %>%
    format_simulation_data() %>%
    magrittr::extract2("p_values") %>%
    calculate_power_t2error(),
  .progress = TRUE
)
```


## Example: Adjust Multiple Parameters

It is perhaps more likely that analysts will want to iterate simulations over a variety of different parameters at once. The code below provides a structure for creating a combination grid based on the 5 input parameters that can be altered; this example can easily be altered to include desired parameters by replacing/removing/expanding the listed parameters. 

Note that the `prob0_list` and `prob1_list` must always be of the same length **and** the corresponding sub-elements of the list must also be of the same length. Put in terms of the application, there must be a Group 2 if there is a Group 1, and the vector representing the number of possible outcomes must be the same length for these 2 groups.

```{r, warning=FALSE}
# Choose sample sizes
sample_size <- c(50, 100)
# Set sample distributions as a proportion c(group1, group2)
sample_prob <- list(c(0.5, 0.5), c(0.75, 0.25))
# Trial 1 has matching probabilities between the 2 groups.
# Trial 2 has non-matching probabilities
prob0_list <- list(
  trial1_group1 = c(0.1, 0.2, 0.3, 0.4),
  trial2_group1 = c(0.1, 0.2, 0.3, 0.4)
)
prob1_list <- list(
  trial1_group2 = c(0.1, 0.2, 0.3, 0.4),
  trial2_group2 = c(0.2, 0.3, 0.3, 0.2)
)
# Number of iterations
niter <- c(20, 100)

# Use tidyr::expand_grid as it creates a tibble,
# which supports the nested tibble structure
info_table <- tidyr::expand_grid(
  sample_size,
  sample_prob,
  prob0_list,
  prob1_list,
  niter
)

info_table

# Calculate either Power/T2 error or T1 error depending on your specific needs
many_sims <- info_table %>%
  purrr::pmap(
    ~ run_simulations(
      sample_size = ..1,
      sample_prob = ..2,
      prob0 = ..3,
      prob1 = ..4,
      niter = ..5
    ) %>%
      format_simulation_data() %>%
      magrittr::extract2("p_values") %>%
      calculate_power_t2error(),
    .progress = TRUE
  )

many_sims[1]
```

