---
title: "Using {ordinalsimr}"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using {ordinalsimr}}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ordinalsimr)
```


# Shiny App

The {ordinalsimr} package wraps a Shiny application and supporting functions for running simulation studies on several pre-selected statistical tests applicable to ordinal data. The Shiny app is particularly suited for calculating Power and Type II error for a proposed 2-group comparison of an ordinal endpoint. Available parameters to manipulate before running the tests include:

* Number of simulations to run (integer)
* The sample size (integer)
* The distribution ratio between group 1 and group 2 (ratio of form X:X)
* Group 1 and Group 2 distribution probabilities for levels of the ordinal outcome

In addition to these parameters for running the simulation, the following can be adjusted in the Distributions page

* Filtering out the top x% of observations from view in the plot. *This does not remove or alter data from any other processes. It is only included as a convenience function for observing smaller windows of the data.*
* Setting the alpha significance level/p-value threshold
* Setting the confidence intervals for the Power/T2 Error calculations
* Setting the confidence intervals (separately) for the Type I Error in Group 1 and Group 2

Bug reports and feature requests can be submitted as issues at [https://github.com/NeuroShepherd/ordinalsimr/issues](https://github.com/NeuroShepherd/ordinalsimr/issues)

## App Options

If you find yourself consistently changing the default parameters of the application to new values manually, it is possible to instead set default values for these parameters. All available preset parameters are prefixed with `ordinalsimr.`, but this functionality has so far only been implemented for a limited number of defaults. Options can be set using the `options()` function. 

Example of changing the default number of rows to 6:

```
options(ordinalsimr.default_entry_rows = 6)
```

### Available Options

* `ordinalsimr.default_entry_rows`: single integer indicating number of rows
* `ordinalsimr.default_distributions`: data frame to use for simulations. Safety checks have not been included on this option. Input must be a 2-column data frame. Column names will be discarded. Recommend that you ensure the sum of values in each column are **near** 1 (e.g. using `dplyr::near()`).


### Planned Options

* `ordinalsimr.default_sample_size`: sample size
* `ordinalsimr.default_simulations`: number of simulation iterations
* ...

## Using Downloaded Data

Data downloaded after running a simulation is stored as a `.rds` file, and can be loaded into your R session using `readRDS()`. The data is structured as a named list with 3 elements at the top level, and several sub-elements. A summary of the available information is available in the code below.

```{r}
output <- readRDS("data-2024-02-16-049dcfcd9afdad1402588ca3a3f73f95-1.rds")

str(output, max.level = 2)

```



# Coding Your Own Simulations

This guide will provide a rough overview of how to code your own simulations using the components of this package should you find the Shiny application too limiting for your own purposes. Key information on functions:

* `run_simulations()` will take simulation input parameters and return a list. The first element contains a matrix of p-values for each run of the simulation, while the second element is multiple lists of information describing group allocations in detail.
* `format_simulation_data()` is recommended for reformatting the two list elements mentioned above. It will return a list of two tibbles.
* `calculate_power_t2error()` and `calculate_t1_error()` can receive the p-value data frames for performing T1 Error, Power, and T2 Error calculations with confidence intervals. See function documentation for additional arguments.

## Example: Power and Type II Error

```{r, warning=FALSE}
sim_results <- run_simulations(sample_size = 80,
                sample_prob = c(0.5,0.5),
                prob0 = c(0.1,0.2,0.3,0.4),
                prob1 = c(0.6,0.2,0.1,0.1),
                niter = 20
                )

formatted_results <- format_simulation_data(sim_results)
names(formatted_results)

formatted_results$p_values

calculate_power_t2error(formatted_results$p_values, alpha = 0.05, power_confidence_int = 95)

```


## Example: Type I Error

To find the Type I error of a distribution, the code from before is largely unchanged except for the fact that the probability vectors set in `run_simulations` must now be equivalent and the `calculate_t1_error()` function is now applied.


```{r, warning=FALSE}
sim_results <- run_simulations(sample_size = 80,
                sample_prob = c(0.5,0.5),
                prob0 = c(0.1,0.2,0.3,0.4),
                prob1 = c(0.1,0.2,0.3,0.4),
                niter = 20
                )

formatted_results <- format_simulation_data(sim_results)
names(formatted_results)

formatted_results$p_values

calculate_t1_error(formatted_results$p_values, alpha = 0.05, t1_error_confidence_int = 95)
```



## Example: Mapping Over Many Sample Sizes

The current version of the application can only accept one sample size at a time. A future iteration of the application will ideally implement this feature, but, in the meantime, there are many situations where it would be advantageous to model test results over many different sample sizes. Depending on how big the number of iterations per sample sizes (`niter`) and the actual number of sample sizes being checked, it may only be practical to do this in a parallelized manner with e.g. {furrr} or {parallel}. In any case, an example of such code is included below:


```{r, warning=FALSE}

sample_sizes <- c(30,50,100)

purrr::map(sample_sizes,
  ~ run_simulations(sample_size = .x,
                sample_prob = c(0.5,0.5),
                prob0 = c(0.1,0.2,0.3,0.4),
                prob1 = c(0.6,0.2,0.1,0.1),
                niter = 100
                ) %>%
    format_simulation_data() %>%
    extract2("p_values") %>%
    calculate_power_t2error()
  )

```





